Picky with peakpicking: assessing chromatographic peak quality with simple

### 2023-07-21 manuscript/manuscript.Rmd
Renaming figures with numbers for upload

```{r Fig1 pred_prob_class_color_hists}
base_gp <- ggplot() +
  aes(x=pred_prob_min, fill=feat_class) +
  facet_wrap(~dataset, scales="free_y") +
  scale_fill_manual(breaks = c("Good", "Bad", "Ambiguous", "Stans only", "Unclassified"),
                    values = c("#f4bb23", "#a41118", "#0b505c", "#028e34", "grey50")) +
  labs(x="Predicted probability", y="# of mass features", fill="Manual\nclassification") +
  theme_bw() +
  theme(axis.text.y.right = element_blank(),
        axis.ticks.y.right = element_blank(),
        legend.position = "none")
axis_breaks <- pred_probs %>%
  group_by(dataset) %>%
  mutate(pred_prob_min=cut(pred_prob_min, breaks=seq(0, 1, 0.02))) %>%
  count(pred_prob_min) %>%
  slice(1) %>%
  pull(n, dataset)
falk_hist_gp <- base_gp + 
  geom_histogram(breaks=seq(0, 1, 0.02), data = pred_probs %>% 
                   filter(dataset=="Falkor")) +
  scale_y_continuous(breaks = c(0, 50, 100, 150, 195),
                     labels = c(0, 50, 100, 150, axis_breaks["Falkor"])) +
  annotate("rect", xmin=0, xmax=1, ymin=180, ymax=185, fill="white", color="black") +
  annotate("rect", xmin=0, xmax=0.125, ymin=195, ymax=Inf, fill="white", color="grey90") +
  coord_cartesian(ylim = c(0, 200), expand = FALSE) +
  theme(axis.title.x = element_blank(), axis.text.x = element_blank())
meso_hist_gp <- base_gp + 
  geom_histogram(breaks=seq(0, 1, 0.02), data = pred_probs %>% 
                   filter(dataset=="MESOSCOPE")) +
  scale_y_continuous(breaks = c(0, 50, 100, 150, 200, 240),
                     labels = c(0, 50, 100, 150, 200, axis_breaks["MESOSCOPE"])) +
  annotate("rect", xmin=0, xmax=1, ymin=226, ymax=232, fill="white", color="black") +
  annotate("rect", xmin=0, xmax=0.125, ymin=240, ymax=Inf, fill="white", color="grey90") +
  coord_cartesian(ylim = c(0, 250), expand = FALSE) +
  theme(axis.title = element_blank(), axis.text.x = element_blank())
culture_hist_gp <- base_gp + 
  geom_histogram(breaks=seq(0, 1, 0.02), data = pred_probs %>% 
                   filter(dataset=="CultureData")) +
  scale_y_continuous(breaks = c(0, 50, 100, 116),
                     labels = c(0, 50, 100, axis_breaks["CultureData"])) +
  geom_vline(xintercept = 0.9, color="black", linewidth=0.5, lty=2) +
  annotate("rect", xmin=0, xmax=1, ymin=110, ymax=113, fill="white", color="black") +
  annotate("rect", xmin=0, xmax=0.125, ymin=116, ymax=Inf, fill="white", color="grey90") +
  coord_cartesian(ylim = c(0, 120), expand = FALSE)
pttime_hist_gp <- base_gp + 
  geom_histogram(breaks=seq(0, 1, 0.02), data = pred_probs %>% 
                   filter(dataset=="Pttime")) +
  scale_y_continuous(breaks = c(0,100,200,300,400,510),
                     labels = c(0,100,200,300,400,axis_breaks["Pttime"])) +
  geom_vline(xintercept = 0.9, color="black", linewidth=0.5, lty=2) +
  annotate("rect", xmin=0, xmax=1, ymin=479, ymax=492, fill="white", color="black") +
  annotate("rect", xmin=0, xmax=0.125, ymin=510, ymax=Inf, fill="white", color="grey90") +
  coord_cartesian(ylim = c(0, 520), expand = FALSE) +
  theme(axis.title.y = element_blank())
egg_grob <- egg::ggarrange(falk_hist_gp, meso_hist_gp, culture_hist_gp, pttime_hist_gp, draw = FALSE)
egg_legend <- cowplot::get_legend(culture_hist_gp + theme(legend.position = "right"))
egg_complete <- cowplot::plot_grid(egg_grob, egg_legend, rel_widths = c(1, 0.25)) +
  theme(plot.background = element_rect(fill = "white", color=NA))
ggsave(plot = egg_complete, 
       filename = "manuscript/figures/pred_prob_class_color_hists.png", 
       device = "png",
       width = 6.5, height = 4, units = "in", dpi = 300)
```

@@

ggsave(plot = fdr_gff_cross_train_gp, 
       filename = "fdr_gff_cross_train.png", 
       device = "png", path = "manuscript/figures",
       width = 6.5, height = 3, units = "in", dpi = 300)
```

@@

raw_prob_space_gp <- dotplot_stepdown_data %>%
  pivot_wider(names_from = training_set, values_from = pred_prob) %>%
  filter(feat_class%in%c("Good", "Bad")) %>%
  mutate(which_params=factor(
    which_params, levels=c("all_params", "xcms_params", "raw_data_params"),
    labels=c("Full model", "XCMS-only model", "Two-parameter model"))) %>%
  ggplot() +
  geom_abline(slope = 1, intercept = 0, color="black", lwd=1) +
  geom_point(aes(x=FT2040_features, y=MS3000_features, color=feat_class), alpha=0.2) +
  geom_text(aes(label=pcor), x=-Inf, y=Inf, hjust=-0.1, vjust=1.5, data=stepdown_cors) +
  facet_wrap(~which_params, nrow = 1) +
  scale_fill_manual(values = c("#a41118", "#f4bb23"), breaks = c("Bad", "Good"), 
                    aesthetics = c("color", "fill")) +
  scale_x_continuous(breaks = c(0, 0.5, 1)) +
  scale_y_continuous(breaks = c(0, 0.5, 1)) +
  coord_fixed() +
  labs(x="Falkor-trained model (predicted probability)", 
       y="MESO-trained model\n(predicted probability)") +
  theme_bw() +
  theme(legend.position = "none")
rank_prob_space_gp <- dotplot_stepdown_data %>%
  group_by(training_set, which_params) %>%
  mutate(pred_prob=rank(pred_prob)) %>%
  pivot_wider(names_from = training_set, values_from = pred_prob) %>%
  mutate(which_params=factor(
    which_params, 
    levels=c("all_params", "xcms_params", "raw_data_params"),
    labels=c("Full model", "XCMS-only model", "Two-parameter model"))) %>%
  filter(feat_class%in%c("Good", "Bad")) %>%
  ggplot() +
  geom_abline(slope = 1, intercept = 0, color="black", lwd=1) +
  geom_point(aes(x=FT2040_features, y=MS3000_features, color=feat_class), alpha=0.2) +
  geom_text(aes(label=scor), x=-Inf, y=Inf, hjust=-0.1, vjust=1.5, data=stepdown_cors) +
  facet_wrap(~which_params, nrow = 1) +
  scale_fill_manual(values = c("#a41118", "#f4bb23"), breaks = c("Bad", "Good"), 
                    aesthetics = c("color", "fill")) +
  scale_x_continuous(breaks = c(0, nrow(MS3000_features)), 
                     labels = c("Min rank", "Max rank")) +
  scale_y_continuous(breaks = c(0, nrow(MS3000_features)),
                     labels = c("Min\nrank", "Max\nrank")) +
  labs(x="Falkor-trained model (ranked likelihood)", 
       y="MESO-trained model\n(ranked likelihood)",
       color=NULL) +
  guides(color=guide_legend(override.aes = list(size=4, alpha=0.8))) +
  coord_fixed() +
  theme_bw() +
  theme(axis.text.x = ggtext::element_markdown(hjust=c(0, 1)),
        # strip.text = element_blank(),
        legend.position=c(0.925, 0.25),
        legend.background = element_blank(),
        legend.spacing.x = unit(0.003, units = "npc"))
comb_prob_space_gp <- egg::ggarrange(raw_prob_space_gp, rank_prob_space_gp, draw = FALSE)
ggsave(plot = comb_prob_space_gp, 
       filename = "comb_prob_space.png", 
       device = "png", path = "manuscript/figures",
       width = 6.5, height = 5, units = "in", dpi = 300)
```

@@

ggsave(plot = fdr_gff_all_datasets_gp, 
       filename = "fdr_gff_all_datasets.png", 
       device = "png", path = "manuscript/figures",
       width = 3, height = 4, units = "in", dpi = 300)
```

@@

clockplot_withboxplots <- gridExtra::arrangeGrob(
  box_list[[1]], clockplot_base_gp, box_list[[2]], 
  layout_matrix = matrix(c(1,2,2,2,3),nrow = 1)
)
clockplot_complete_gp <- gridExtra::arrangeGrob(
  clockplot_hist_gp, clockplot_withboxplots, 
  layout_matrix = matrix(c(1,2,2), ncol=1)
)
ggsave(plot = clockplot_complete_gp, 
       filename = "clockplot_complete.png", 
       device = "png", path = "manuscript/figures",
       width = 6.5, height = 7, units = "in", dpi = 300)
```

@@

```{r Fig6 NMDS results}
thresh_NMDS_gp <- thresh_NMDS_df %>%
  ggplot() +
  geom_vline(xintercept = 0) +
  geom_hline(yintercept = 0) +
  geom_polygon(aes(x=MDS1, y=MDS2, fill=depth, color=depth, group=basename), alpha=0.5) +
  geom_label(aes(label=stress), x=Inf, y=Inf,  label.r = unit(0, "in"), 
             hjust=1, vjust=1, data=distinct(thresh_NMDS_df, mat_name, stress)) +
  facet_wrap(~mat_name, scales = "free", ncol=3) +
  scale_color_manual(values=viridis::viridis(9, direction = -1)[c(2,5,8)], 
                     breaks = c("15m", "DCM", "175m"),
                     aesthetics = c("color", "fill"),
                     name="Sample\ndepth") +
  theme_bw() +
  theme(aspect.ratio = 1)
ggsave(plot = thresh_NMDS_gp, filename = "thresh_NMDS_plots.png", 
       device = "png", path = "manuscript/figures",
       width = 6.5, height = 4, units = "in", dpi = 300)
```

@@

comb_cowplot <- cowplot::plot_grid(good_chroms_gp, bad_chroms_gp, chroms_legend, 
                                   ncol = 1, rel_heights = c(1, 1, 0.15))
ggsave(plot = comb_cowplot, 
       filename = "somegood_somebad_depthcolor_chroms.png", 
       device = "png", path = "manuscript/figures",
       width = 6.5, height = 3, units = "in", dpi = 300)
```

@@

ggsave(plot = comb_peakmetrics_gp, 
       filename = "peakmetrics_singlechrom.png", 
       device = "png", path = "manuscript/figures",
       width = 6.5, height = 4, units = "in", dpi = 300)
```

>>>>

```{r Fig1 pred_prob_class_color_hists}
base_gp <- ggplot() +
  aes(x=pred_prob_min, fill=feat_class) +
  facet_wrap(~dataset, scales="free_y") +
  scale_fill_manual(breaks = c("Good", "Bad", "Ambiguous", "Stans only", "Unclassified"),
                    values = c("#f4bb23", "#a41118", "#0b505c", "#028e34", "grey50")) +
  labs(x="Predicted probability", y="# of mass features", fill="Manual\nclassification") +
  theme_bw() +
  theme(axis.text.y.right = element_blank(),
        axis.ticks.y.right = element_blank(),
        legend.position = "none")
axis_breaks <- pred_probs %>%
  group_by(dataset) %>%
  mutate(pred_prob_min=cut(pred_prob_min, breaks=seq(0, 1, 0.02))) %>%
  count(pred_prob_min) %>%
  slice(1) %>%
  pull(n, dataset)
falk_hist_gp <- base_gp + 
  geom_histogram(breaks=seq(0, 1, 0.02), data = pred_probs %>% 
                   filter(dataset=="Falkor")) +
  scale_y_continuous(breaks = c(0, 50, 100, 150, 195),
                     labels = c(0, 50, 100, 150, axis_breaks["Falkor"])) +
  annotate("rect", xmin=0, xmax=1, ymin=180, ymax=185, fill="white", color="black") +
  annotate("rect", xmin=0, xmax=0.125, ymin=195, ymax=Inf, fill="white", color="grey90") +
  coord_cartesian(ylim = c(0, 200), expand = FALSE) +
  theme(axis.title.x = element_blank(), axis.text.x = element_blank())
meso_hist_gp <- base_gp + 
  geom_histogram(breaks=seq(0, 1, 0.02), data = pred_probs %>% 
                   filter(dataset=="MESOSCOPE")) +
  scale_y_continuous(breaks = c(0, 50, 100, 150, 200, 240),
                     labels = c(0, 50, 100, 150, 200, axis_breaks["MESOSCOPE"])) +
  annotate("rect", xmin=0, xmax=1, ymin=226, ymax=232, fill="white", color="black") +
  annotate("rect", xmin=0, xmax=0.125, ymin=240, ymax=Inf, fill="white", color="grey90") +
  coord_cartesian(ylim = c(0, 250), expand = FALSE) +
  theme(axis.title = element_blank(), axis.text.x = element_blank())
culture_hist_gp <- base_gp + 
  geom_histogram(breaks=seq(0, 1, 0.02), data = pred_probs %>% 
                   filter(dataset=="CultureData")) +
  scale_y_continuous(breaks = c(0, 50, 100, 116),
                     labels = c(0, 50, 100, axis_breaks["CultureData"])) +
  geom_vline(xintercept = 0.9, color="black", linewidth=0.5, lty=2) +
  annotate("rect", xmin=0, xmax=1, ymin=110, ymax=113, fill="white", color="black") +
  annotate("rect", xmin=0, xmax=0.125, ymin=116, ymax=Inf, fill="white", color="grey90") +
  coord_cartesian(ylim = c(0, 120), expand = FALSE)
pttime_hist_gp <- base_gp + 
  geom_histogram(breaks=seq(0, 1, 0.02), data = pred_probs %>% 
                   filter(dataset=="Pttime")) +
  scale_y_continuous(breaks = c(0,100,200,300,400,510),
                     labels = c(0,100,200,300,400,axis_breaks["Pttime"])) +
  geom_vline(xintercept = 0.9, color="black", linewidth=0.5, lty=2) +
  annotate("rect", xmin=0, xmax=1, ymin=479, ymax=492, fill="white", color="black") +
  annotate("rect", xmin=0, xmax=0.125, ymin=510, ymax=Inf, fill="white", color="grey90") +
  coord_cartesian(ylim = c(0, 520), expand = FALSE) +
  theme(axis.title.y = element_blank())
egg_grob <- egg::ggarrange(falk_hist_gp, meso_hist_gp, culture_hist_gp, pttime_hist_gp, draw = FALSE)
egg_legend <- cowplot::get_legend(culture_hist_gp + theme(legend.position = "right"))
egg_complete <- cowplot::plot_grid(egg_grob, egg_legend, rel_widths = c(1, 0.25)) +
  theme(plot.background = element_rect(fill = "white", color=NA))
ggsave(plot = egg_complete, 
       filename = "manuscript/figures/fig_1_pred_prob_class_color_hists.png", 
       device = "png",
       width = 6.5, height = 4, units = "in", dpi = 300)
```

@@

ggsave(plot = fdr_gff_cross_train_gp, 
       filename = "fig_2_fdr_gff_cross_train.png", 
       device = "png", path = "manuscript/figures",
       width = 6.5, height = 3, units = "in", dpi = 300)
```

@@

raw_prob_space_gp <- dotplot_stepdown_data %>%
  pivot_wider(names_from = training_set, values_from = pred_prob) %>%
  filter(feat_class%in%c("Good", "Bad")) %>%
  mutate(which_params=factor(
    which_params, levels=c("all_params", "xcms_params", "raw_data_params"),
    labels=c("Full model", "XCMS-only model", "Two-parameter model"))) %>%
  ggplot() +
  geom_abline(slope = 1, intercept = 0, color="black", lwd=1) +
  geom_point(aes(x=FT2040_features, y=MS3000_features, color=feat_class), alpha=0.2) +
  geom_text(aes(label=pcor), x=-Inf, y=Inf, hjust=-0.1, vjust=1.5, data=stepdown_cors) +
  facet_wrap(~which_params, nrow = 1) +
  scale_fill_manual(values = c("#a41118", "#f4bb23"), breaks = c("Bad", "Good"), 
                    aesthetics = c("color", "fill")) +
  scale_x_continuous(breaks = c(0, 0.5, 1)) +
  scale_y_continuous(breaks = c(0, 0.5, 1)) +
  coord_fixed() +
  labs(x="Falkor-trained model (predicted probability)", 
       y="MESO-trained model\n(predicted probability)") +
  theme_bw() +
  theme(legend.position = "none")
rank_prob_space_gp <- dotplot_stepdown_data %>%
  group_by(training_set, which_params) %>%
  mutate(pred_prob=rank(pred_prob)) %>%
  pivot_wider(names_from = training_set, values_from = pred_prob) %>%
  mutate(which_params=factor(
    which_params, 
    levels=c("all_params", "xcms_params", "raw_data_params"),
    labels=c("Full model", "XCMS-only model", "Two-parameter model"))) %>%
  filter(feat_class%in%c("Good", "Bad")) %>%
  ggplot() +
  geom_abline(slope = 1, intercept = 0, color="black", lwd=1) +
  geom_point(aes(x=FT2040_features, y=MS3000_features, color=feat_class), alpha=0.2) +
  geom_text(aes(label=scor), x=-Inf, y=Inf, hjust=-0.1, vjust=1.5, data=stepdown_cors) +
  facet_wrap(~which_params, nrow = 1) +
  scale_fill_manual(values = c("#a41118", "#f4bb23"), breaks = c("Bad", "Good"), 
                    aesthetics = c("color", "fill")) +
  scale_x_continuous(breaks = c(0, nrow(MS3000_features)), 
                     labels = c("Min rank", "Max rank")) +
  scale_y_continuous(breaks = c(0, nrow(MS3000_features)),
                     labels = c("Min\nrank", "Max\nrank")) +
  labs(x="Falkor-trained model (ranked likelihood)", 
       y="MESO-trained model\n(ranked likelihood)",
       color=NULL) +
  guides(color=guide_legend(override.aes = list(size=4, alpha=0.8))) +
  coord_fixed() +
  theme_bw() +
  theme(axis.text.x = ggtext::element_markdown(hjust=c(0, 1)),
        # strip.text = element_blank(),
        legend.position=c(0.925, 0.25),
        legend.background = element_blank(),
        legend.spacing.x = unit(0.003, units = "npc"))
comb_prob_space_gp <- egg::ggarrange(raw_prob_space_gp, rank_prob_space_gp, draw = FALSE)
ggsave(plot = comb_prob_space_gp, 
       filename = "fig_3_comb_prob_space.png", 
       device = "png", path = "manuscript/figures",
       width = 6.5, height = 5, units = "in", dpi = 300)
```

@@

ggsave(plot = fdr_gff_all_datasets_gp, 
       filename = "fig_4_fdr_gff_all_datasets.png", 
       device = "png", path = "manuscript/figures",
       width = 3, height = 4, units = "in", dpi = 300)
```

@@

clockplot_withboxplots <- gridExtra::arrangeGrob(
  box_list[[1]], clockplot_base_gp, box_list[[2]], 
  layout_matrix = matrix(c(1,2,2,2,3),nrow = 1)
)
clockplot_complete_gp <- gridExtra::arrangeGrob(
  clockplot_hist_gp, clockplot_withboxplots, 
  layout_matrix = matrix(c(1,2,2), ncol=1)
)
ggsave(plot = clockplot_complete_gp, 
       filename = "fig_5_clockplot_complete.png", 
       device = "png", path = "manuscript/figures",
       width = 6.5, height = 7, units = "in", dpi = 300)
```

@@

```{r Fig6 NMDS results}
thresh_NMDS_gp <- thresh_NMDS_df %>%
  ggplot() +
  geom_vline(xintercept = 0) +
  geom_hline(yintercept = 0) +
  geom_polygon(aes(x=MDS1, y=MDS2, fill=depth, color=depth, group=basename), alpha=0.5) +
  geom_label(aes(label=stress), x=Inf, y=Inf,  label.r = unit(0, "in"), 
             hjust=1, vjust=1, data=distinct(thresh_NMDS_df, mat_name, stress)) +
  facet_wrap(~mat_name, scales = "free", ncol=3) +
  scale_color_manual(values=viridis::viridis(9, direction = -1)[c(2,5,8)], 
                     breaks = c("15m", "DCM", "175m"),
                     aesthetics = c("color", "fill"),
                     name="Sample\ndepth") +
  theme_bw() +
  theme(aspect.ratio = 1)
ggsave(plot = thresh_NMDS_gp, filename = "fig_6_thresh_NMDS_plots.png", 
       device = "png", path = "manuscript/figures",
       width = 6.5, height = 4, units = "in", dpi = 300)
```

@@

comb_cowplot <- cowplot::plot_grid(good_chroms_gp, bad_chroms_gp, chroms_legend, 
                                   ncol = 1, rel_heights = c(1, 1, 0.15))
ggsave(plot = comb_cowplot, 
       filename = "fig_7_somegood_somebad_depthcolor_chroms.png", 
       device = "png", path = "manuscript/figures",
       width = 6.5, height = 3, units = "in", dpi = 300)
```

@@

ggsave(plot = comb_peakmetrics_gp, 
       filename = "fig_8_peakmetrics_singlechrom.png", 
       device = "png", path = "manuscript/figures",
       width = 6.5, height = 4, units = "in", dpi = 300)
```
